White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions

### 2017-09-19 papers/wdmerger_II/paper.tex
Remove nuclear network tests; these are outside the scope

White dwarfs are mainly composed of $\alpha$-chain particles, primarily ${}^4$He,
${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, and ${}^{24}$ Mg. Therefore the core of
any network appropriate for modeling nuclear burning in white dwarfs will be
these alpha chain nuclides, with the idea being that links up the $\alpha$-chain
will eventually get us to ${}^{56}$Ni, the nuclide responsible for the
energy output of Type Ia supernovae. In this paper we consider four networks
to do this, presented in order of increasing complexity. The most simple is
\isoseven\ \citep{timmes:2000}, which includes all of the aforementioned isotopes and
${}^{28}$Si (see also \citet{hix:1998}). ${}^{28}$Si effectively measures the
equilibrium state of silicon-group elements, and ${}^{56}$Ni effectively measures
the equilibrium state of iron-group elements, with the link between them governed
by the effective loss or gain of seven $\alpha$-particles. This type of network
was used by \citet{rosswog:2009} for their collision calculations in SPH.

Next is \aproxthirteen\ \citep{timmes:1999,timmes:2000}. This includes
all of the isotopes of \isoseven, and all of the $\alpha$-chain particles between
silicon and nickel (${}^{32}$S, ${}^{36}$Ar, ${}^{40}$Ca, ${}^{44}$Ti, ${}^{48}$Cr,
and ${}^{52}$Fe). This network was used by \citet{hawley:2012} and \citet{raskin:2010}.
\citet{loren-aguilar:2010} and \citet{garcia-senz:2013} used a very similar network
that included additionally ${}^{60}$Zn. The \aproxnineteen\ network \citep{timmes:1999}
builds on \aproxthirteen\ by including isotopes for hydrogen burning and explicit
tracking of photodisintegration into ${}^{54}$Fe. This network was used by
\citet{kushnir:2013}, \citet{kushnir:2014}, and \citet{rosswog:2009} for their
calculations with FLASH, and \citet{papish:2015}. Finally we will also
consider \aproxtwentyone, which includes all of the above plus ${}^{56}$Cr
and ${}^{56}$Fe and related reaction links. The primary virtue of using
the latter two networks is that they allow us to track changes away from
$Y_e = 0.5$.

All four of these networks have been ported into a form that is consistent
with the \amrex\ codes, in the freely available \microphysics\ code
repository\footnote{\microphysics\ can be obtained at \url{https://github.com/starkiller-astro/Microphysics}.},
a collection of microphysical routines that are designed to be used in our
hydrodynamics codes. These can be easily swapped at compile time by using the 
appropriate makefile variable.

@@

\subsection{Nuclear Network}
\label{sec:parameters:network}

Now we move to a discussion of the effect of including various isotopes
in the nuclear reaction network. In particular we compare the \isoseven, \aproxthirteen,
\aproxnineteen, and \aproxtwentyone\ reaction networks. In all cases we run
the same problem setup, and the unused isotopes (everything but carbon and oxygen)
have their mass fractions in every zone set to a negligibly small number at initialization.
\autoref{table:networks} lists the $^{56}$Ni generation and total energy generation (determined
by subtracting the initial energy from the final energy). The energy generated in the latter
three cases agrees remarkably well, with a less than $0.5\%$ difference. \aproxthirteen\
over-generates nickel at the $2\%$ level compared to the networks with more isotopes, a reasonably
small difference. On the other hand, agreement with \isoseven\ is not very strong, as \isoseven\
generates far too much nickel and far too little energy. Nevertheless while the quantitative
agreement is not there, the collision looks qualitatively similar compared to the evolution with
more complicated networks, so while \isoseven\ should not be used for predicting nucleosynthetic
yields in production science runs, it may still be fair for use in proof-of-concept studies.

\input{plots/networks_m_P_0.64_m_S_0.64.tbl}

\isoseven\ also has the peculiar property that after the detonation has passed through the WDs,
the region at the central contact point does not fully stop burning, yielding a quasistatic energy
release that can add thousands to millions of extra timesteps to the simulation (in addition to the few
hundred it normally takes, using our designated stopping criterion). The full effect of these additional
timesteps only amounts to a trace change in the nickel production. By examining the behavior of zones
near the collision point, we discovered the reason for this. The zones in question have a composition that
is approximately $98\%$ Ni and $2\%$ He by mass. For this combination, a direct evaluation of the RHS in
comparison to, say, \aproxthirteen, reveals that $dX/dt$ for these isotopes is off by many orders of magnitude.
This is a consequence of the choice inherent to the network to assume that the isotopes in between $^{28}$Si
and $^{56}$Ni are in an effective equilibrium state. So an advective update that yields even a small change
to the abundances in the zone will very strongly knock the zone out of equilibrium with respect to the burn
step following the advective update. This then causes the burning timestep limiter to kick in and make the
timestep very small so that the advective changes do not cause such strong effects on the burning equilibrium.
This effect is an argument against the use of \isoseven\ for explosive, non-hydrostatic burning problems.

\aproxnineteen\ and \aproxtwentyone\ both take significantly longer to evaluate at our default
ODE tolerances than \aproxthirteen\ does, even after accounting for the increased number of species.
This is due to a significant number of integration failures and subsequent retries (see the discussion
in \autoref{sec:burner}). The species tolerance of $10^{-8}$ is challenging for these networks.

>>>>

White dwarfs are mainly composed of $\alpha$-chain particles, primarily ${}^4$He,
${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, and ${}^{24}$ Mg. Therefore the core of
any network appropriate for modeling nuclear burning in white dwarfs will be
these alpha chain nuclides, with the idea being that links up the $\alpha$-chain
will eventually get us to ${}^{56}$Ni, the nuclide responsible for the
energy output of Type Ia supernovae. In this paper we use the \aproxthirteen\
network\ \citep{timmes:1999,timmes:2000}. This includes all thirteen $\alpha$-chain
particles between ${}^4$He and ${}^{56}$Ni. This network was used by \citet{hawley:2012}
and \citet{raskin:2010}. \citet{loren-aguilar:2010} and \citet{garcia-senz:2013} used a
very similar network that additionally included ${}^{60}$Zn. This network has been
ported into a form that is consistent with the \amrex\ codes, in the freely available
\microphysics\ code repository
\footnote{\microphysics\ can be obtained at \url{https://github.com/starkiller-astro/Microphysics}.},
a collection of microphysical routines that are designed to be used in our
hydrodynamics codes. In our \microphysics\ repository we have also ported \aproxnineteen\
\citep{timmes:1999}, which was used in a few previous collision papers, and \aproxtwentyone.
The isotopes added by these networks relative to \aproxthirteen\ are not particularly relevant
to the problem of head-on white dwarf collisions, so we do not discuss results with these networks
here (though we have verified for this problem that these larger networks do not significantly change
the answer). We also make available \isoseven\ \citep{timmes:2000}, which skips the nuclides between
silicon and nickel, and was used by \citet{rosswog:2009}. \isoseven\ gives significantly different
quantitative results than \aproxthirteen\ for this problem, due to the inapplicable assumption of
an equilibrium abundance of the unused isotopes, but the results are still qualitatively similar.

@@
